---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ecoteach

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ecoteach)](https://CRAN.R-project.org/package=ecoteach)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

The goal of ecoteach is to provide a collection of curated educational datasets for teaching ecology and agriculture concepts. The package includes data on wildlife monitoring, plant treatments, and ecological observations with documentation and examples for educational use.

## Installation

You can install the released version of ecoteach from [CRAN](https://CRAN.R-project.org) with:

```{r, eval = FALSE}
install.packages("ecoteach")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("username/ecoteach")
```

## Example

This package provides datasets that can be used to teach ecological and statistical concepts. Here's an example using the dormouse hibernation dataset:

```{r example}
# Load the package
library(ecoteach)
# Or use devtools::load_all() during development
# devtools::load_all()

# Load the dormouse hibernation dataset
data(dormouse_hibernation)

# View the first few rows
head(dormouse_hibernation)
```

### Exploring the data

The dormouse hibernation dataset contains data on hibernation patterns and body mass changes in edible dormice (*Glis glis*). We can create a simple visualization to explore how body mass changes during hibernation:

```{r visualization, eval=FALSE}
# Basic analysis of body mass loss during hibernation
library(ggplot2)

# Calculate percentage of body mass lost
dormouse_hibernation$mass_loss_percent <- 
  (dormouse_hibernation$body_mass_before - dormouse_hibernation$body_mass_after) / 
  dormouse_hibernation$body_mass_before * 100

# Create a boxplot of mass loss percentage by sex
ggplot(dormouse_hibernation, aes(x = sex, y = mass_loss_percent, fill = sex)) +
  geom_boxplot() +
  labs(title = "Body Mass Loss During Hibernation in Dormice",
       x = "Sex",
       y = "Body Mass Loss (%)") +
  theme_minimal()
```

This example shows how the dataset can be used to explore sex differences in hibernation physiology, which could be used in a teaching context to demonstrate boxplots, t-tests, or ANOVA.

## Available Datasets

The package includes the following datasets:

```{r datasets, echo=FALSE}
# List all datasets in the package
library(ecoteach)
# Get all objects in the package
datasets <- ls("package:ecoteach")
cat(paste0("- `", datasets, "`\n"))
```

Each dataset includes comprehensive documentation with details about the variables, source references, and suggested analyses for teaching purposes.
