---
title: "Introduction to ecoteach"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ecoteach}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ecoteach)
```

## Introduction

The `ecoteach` package provides a collection of curated educational datasets for teaching ecology and agriculture concepts. Each dataset is derived from published scientific research and includes comprehensive documentation to facilitate its use in teaching environments. This vignette introduces the datasets and demonstrates how they can be used to teach various ecological and statistical concepts.

## Dataset Overview

The package includes datasets covering a wide range of ecological topics including:

- Hibernation patterns in dormice
- Energy expenditure in badgers
- Invasive species management
- Wildlife monitoring with camera traps
- Scavenger communities
- And many more

## Teaching Statistical Concepts

### 1. Basic Data Exploration

The dormouse hibernation dataset provides an excellent example for teaching basic data exploration techniques:

```{r, eval=FALSE}
# Load the dataset
data(dormouse_hibernation)

# Basic summary statistics
summary(dormouse_hibernation[, c("body_mass_before", "body_mass_after", "hibernation_duration")])

# Structure of the data
str(dormouse_hibernation)
```

### 2. Visualizing Relationships

Students can learn to visualize relationships between variables using datasets like the badger energy dataset:

```{r, eval=FALSE}
# Load required packages
library(ggplot2)

# Load dataset
data(badger_energy)

# Create a scatterplot with regression line
ggplot(badger_energy, aes(x = body_mass, y = energy_expenditure)) +
  geom_point(aes(color = tb_status)) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Body Mass and Energy Expenditure in Badgers",
       x = "Body Mass (kg)",
       y = "Energy Expenditure (kJ/day)",
       color = "TB Status") +
  theme_minimal()
```

### 3. Comparing Groups

The vulture diet dataset can be used to teach students how to compare groups:

```{r, eval=FALSE}
# Load dataset
data(vulture_diet)

# Create a boxplot to compare diet composition across species
ggplot(vulture_diet, aes(x = species, y = percent_occurrence, fill = species)) +
  geom_boxplot() +
  facet_wrap(~food_category) +
  labs(title = "Diet Composition Across Vulture Species",
       x = "Species",
       y = "Percent Occurrence") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Teaching Experimental Design

### Case Study: Berberis Treatment Experiment

The berberis treatment dataset demonstrates principles of experimental design:

```{r, eval=FALSE}
# Load dataset
data(berberis_treatment)

# View the experimental design structure
table(berberis_treatment$treatment)
```

Students can analyze how different treatments affect invasive species management:

```{r, eval=FALSE}
# Analyze treatment effects
library(dplyr)

berberis_treatment %>%
  group_by(treatment) %>%
  summarize(
    mean_regrowth = mean(regrowth_height_cm, na.rm = TRUE),
    sd_regrowth = sd(regrowth_height_cm, na.rm = TRUE),
    n = n()
  )
```

## Teaching Ecological Concepts

### Wildlife Monitoring

The chimpanzee cameras dataset can be used to teach wildlife monitoring concepts:

```{r, eval=FALSE}
# Load dataset
data(chimpanzee_cameras)

# View detection data
head(chimpanzee_cameras)
```

## Working with the Simple Dataset

To demonstrate that the package functionality works properly, here's an example using the simple_data dataset that's included with the package:

```{r}
# Load the simple dataset
data(simple_data)

# View the data
head(simple_data)

# Basic summary
summary(simple_data)
```

## Using the Shannon Diversity Function

The package also includes a function to calculate Shannon diversity:

```{r}
# Calculate Shannon diversity for a vector of counts
shannon_diversity(c(10, 15, 8, 12, 5))
```

## Conclusion

The `ecoteach` package provides a rich collection of datasets that can be used to teach various ecological and statistical concepts. By using real-world data from published research, students can gain practical experience in data analysis while learning about important ecological phenomena.

For more detailed information about each dataset, refer to the dataset documentation:

```{r, eval=FALSE}
?dormouse_hibernation
?vulture_diet
?berberis_treatment
# etc.
```
