---
title: "Introduction to ecoteach"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ecoteach}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ecoteach)
```

## Introduction

The `ecoteach` package provides a collection of curated educational datasets for teaching ecology and agriculture concepts. Each dataset is derived from published scientific research and includes comprehensive documentation to facilitate its use in teaching environments.

## Available Datasets

The package includes the following datasets:

- `badger_energy`: Energy expenditure data for European badgers with tuberculosis
- `barnswallow_brightness`: Plumage brightness data for barn swallows
- `berberis_treatment`: Data on invasive Berberis management treatments
- `carrion_arrivals`: Vertebrate scavenger visits to roe deer carrion
- `chimpanzee_cameras`: Camera trap detection data for wild chimpanzees
- `dormouse_hibernation`: Hibernation and reproduction data for edible dormice
- `Dsimulans_matechoice`: Mate copying data for fruit flies
- `elephant_farmers`: Agricultural use metrics for elephants
- `lion_reproduction`: Reproductive data for Galapagos sea lions
- `magellanic_penguins`: Long-term monitoring data on Magellanic penguins
- `pangolin_habitat`: Habitat occupancy data for Chinese pangolins
- `raccoondog_environment`: Raccoon dog activity and environmental factors in China
- `scavenger_community`: Scavenger community structure along environmental gradients
- `vulture_diet`: Diet composition analysis of African vultures
- `whale_brains`: Brain size evolution data for cetaceans

## Using the Datasets

To use any of the datasets in the package, load them with the `data()` function:

```{r}
# Example with dormouse hibernation dataset
data(dormouse_hibernation)
head(dormouse_hibernation)

# Example with vulture diet dataset
data(vulture_diet)
head(vulture_diet)
```

For more detailed information about each dataset, refer to the dataset documentation:

```{r}
# E.G.:
# ?dormouse_hibernation
# ?vulture_diet
# ?berberis_treatment
# etc.
```
